---
title: "Recovering a True Model"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
p1Data <- read.csv("/Users/soren.laney/Desktop/BYUI/Fall Semester 2022/Math 425 Applied Linear Regression/Class Notebooks/p1Data-1.csv", header=TRUE)
p2Data <- read.csv("/Users/soren.laney/Desktop/BYUI/Fall Semester 2022/Math 425 Applied Linear Regression/Class Notebooks/p2Data-1.csv", header=TRUE)
p3Data <- read.csv("/Users/soren.laney/Desktop/BYUI/Fall Semester 2022/Math 425 Applied Linear Regression/Class Notebooks/p3Data-1.csv", header=TRUE)
```

## Part 1 - Recovering a True Model (Beginner Level)

We believe the true model to be...

$$
  Y_i = \underbrace{/beta_0 + /beta_1X_4i +beta_2}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = ...
$$

with our estimate of $\sigma$ as ...

```{r}
pairs(p1Data, panel=panel.smooth)


lm1 <- lm(Y ~ X4, data=p1Data)
summary(lm1)
plot(lm1, which=1)


pairs(cbind(Res = lm1$residuals, p1Data), panel=panel.smooth, col=as.factor(p1Data$X2))



lm2 <- lm(Y ~ X4 + X2, data=p1Data)
summary(lm2)


plot(lm2, which=1)
b <- coef(lm2)
plot(Y~ X4, data=p1Data, col=as.factor(X2))
X2=0
curve(b[1] + b[2]*X4 + b[3]*X2, xname="X4", col="black", add=TRUE)
X2=1
curve(b[1] + b[2]*X4 + b[3]*X2, xname="X4", col="red", add=TRUE)

lm3 <- lm(Y ~ X4 +X2 + X4:X2, data=p1Data)
summary(lm3)
b <- coef(lm3)
plot(Y~ X4, data=p1Data, col=as.factor(X2))
X2=0
curve(b[1] + b[2]*X4 + b[3]*X2, xname="X4", col="black", add=TRUE)
X2=1
curve(b[1] + b[2]*X4 + b[3]*X2, xname="X4", col="red", add=TRUE)

plot(lm3, which=1)

```


## Part 2 - Recovering a True Model (Intermediate Level)

We believe the true model to be...

$$
  Y_i = \underbrace{...}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = ...
$$

with our estimate of $\sigma$ as ...

```{r}
#View(p2Data)
pairs(p2Data)

lm1 <- lm(Y ~ X5 +I(X5^2), data=p2Data)
summary(lm1)

plot(lm1, which=1)


pairs(cbind(Res = lm1$res, p2Data), panel=panel.smooth)

lm2 <- lm(Y ~ X5 + I(X5^2) + X3, data=p2Data)
summary(lm2)


plot(lm2, which=1)


```



## Part 3 - Recovering a True Model (Advanced Level)

We believe the true model to be...

$$
  Y_i = \underbrace{...}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = ...
$$

with our estimate of $\sigma$ as ...

```{r}

```
